---
title: "random forests h2b"
author: "Elena C"
date: "2023-11-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Random forests

Random Forest reduces overfitting by averaging multiple decision trees and is less sensitive to noise and outliers in the data. 
Here we use classification random forests

## Load packages
```{r, message=F}
rm(list = ls())
library(randomForest)
library(datasets)
library(caret)
library(ggplot2)
library(cli)
```

## Hypothesis and data

H2b states: *"There are effects of image novelty (i.e., between images shown for the first time/new vs. repeated/old images) within the time-range from 300–500 ms on theta power at fronto-central channels"*

The data table included in the analysis contains info on the software, high-pass cutoff values, reference, sampling rate, number of excluded subjects, tf baseline window length, number of channels included in the analysis, frequency range that was included in the analysis, whether teams averaged over the time points/channels/frequencies, if ICA component selection was manual or automatic, multiple comparisons method, statistical method, and whether the finding was significant or not for the h2b testing. 


## Load data and inspect it for typos
```{r}

myData_h2b <-read.csv("/Users/ecesnaite/Desktop/BuschLab/EEGManyPipelines/git/EEGManyPipelines-personal/Matlab/Big Analysis/independent_variables_for_models_h2b_complete.csv")

myData_h2b$software[myData_h2b$software=='EEGLAB'] <- 'eeglab'
myData_h2b$reref[myData_h2b$reref == ''] <- 'original'
myData_h2b$hf_cutoff[is.na(myData_h2b$hf_cutoff)] <- 0
myData_h2b$tf_baseline_window[is.na(myData_h2b$tf_baseline_window)] <- 0
myData_h2b$ds_fs[is.na(myData_h2b$ds_fs)] <- 512
myData_h2b$subj_excluded[is.na(myData_h2b$subj_excluded)] <- 0 
myData_h2b$nr_chan_h2b_2[is.na(myData_h2b$nr_chan_h2b_2)] <- mean((myData_h2b$nr_chan_h2b_2), na.rm = T)
myData_h2b$time_window_interest_h2b[is.na(myData_h2b$time_window_interest_h2b)] <- mean((myData_h2b$time_window_interest_h2b), na.rm = T)
myData_h2b$stat_method_h2b[myData_h2b$stat_method_h2b== ""] <- 'unknown'

myData_h2b$comp_selection[myData_h2b$comp_selection == 0] <- 'unknown'
myData_h2b$comp_selection[myData_h2b$comp_selection == 1] <- 'visual'
myData_h2b$comp_selection[myData_h2b$comp_selection == 2] <- 'plugin'

myData_h2b$mc_method_h2b[which(myData_h2b$mc_method_h2b == 'none' & myData_h2b$mc_binary==1)] <- 'unknown'
myData_h2b$mc_method_h2b[which(myData_h2b$mc_method_h2b != 'none' & myData_h2b$mc_binary==0)] <- 'none'
myData_h2b$mc_method_h2b[which(myData_h2b$mc_method_h2b == "" & myData_h2b$mc_binary==1)] <- 'unknown'
myData_h2b$tf_method[myData_h2b$tf_method == ''] <- 'unknown'

data_incl <- myData_h2b[,c(-7, -11)] 

data_incl$sig_res <- as.factor(data_incl$sig_res)
table(data_incl$sig_res) 

data_incl <- na.omit(data_incl)

```

Each node in a Random Forest is not comparing feature values, it is simply splitting a sorted list that requires absolute values for branching. The algorithm is based on partitioning the data to make predictions, therefore, it does not require normalization.

## Data partition into the training and test data sets
```{r}
#Lets start with random seed so the outcome will be repeatable and store train and test data.
set.seed(22)
ind <- sample(2, nrow(data_incl), replace = TRUE, prob = c(0.7, 0.3))
train <- data_incl[ind==1,]
test <- data_incl[ind==2,]
```


## Build a random forest

Here we have created 500 trees, therefore we have got 500 bootstrap samples with us, one for each tree. These samples would consist of m(sqrt(no. of features)) features taken at a time from the data and creates a tree with them.For each bootstrap sample, there is one third of data which was not used in the creation of the tree, i.e., it was out of the sample. This data is referred to as out of bag data. In order to get an unbiased measure of the accuracy of the model over test data, out of bag error is used. The out of bag data is passed for each tree is passed through that tree and the outputs are aggregated to give out of bag error. This percentage error is quite effective in estimating the error in the testing set and does not require further cross validation.

Two parameters should be considered:
-Number of trees used in the forest (ntree)
-Number of random variables used in each tree (mtry)

First set the mtry to the default value (sqrt of total number of all predictors) and search for the optimal ntree value.

## Find the optimal mtry value
```{r}
mtry <- tuneRF(data_incl[,-9],data_incl$sig_res, ntreeTry=500,
               stepFactor=1.5,improve=0.01, trace=TRUE, plot=TRUE)
best.m <- mtry[mtry[, 2] == min(mtry[, 2]), 1]
print(mtry)
print(best.m)
```

## Run the model
```{r}
rf <- randomForest(sig_res~., data=train, proximity=TRUE, mtry=6,ntree=500, importance=T) 
print(rf)
```
The model predicts the data with the error rate of % (or % correct predictions).

## Evaluate variable importance
Mean Decrease Accuracy - How much the model accuracy decreases if we drop that variable.
Mean Decrease Gini - Measure of variable importance based on the Gini impurity index used for the calculation of splits in trees.
```{r}
importance(rf)
varImpPlot(rf)
```

## Dimension plot 

```{r}
MDSplot(rf, train$sig_res)
```
## Error rate of random forests. 
```{r}
#'out-of-bag' (OOB) error  can be used as the validation sample for the fit to assess the regression accuracy for new observations not included in the training data. 
plot(rf)
library(ggplot2)
library(tidyr)
plotdf <- pivot_longer(data.frame(ntrees=1:nrow(rf$err.rate),rf$err.rate),-ntrees)
ggplot(plotdf,aes(x=ntrees,y=value,col=name)) + 
geom_line() + theme_bw()
```

## Validation
```{r}
#1: looking at confusion matrix
prediction <- predict(rf,test[,-9])
table(observed = test[,9], predicted = prediction)
```


