---
title: "AIC H1"
author: "Elena C"
date: "2024-01-10"
output: html_document
---

## Load packages
```{r setup, warning=FALSE}
library(tidyverse)
library(broom)
library(readxl)
library(AICcmodavg)
library(ggplot2)
```

## Load data and inspect it for typos


```{r}
h1_data <- read.csv("C:/Users/ecesnait/Desktop/EEGManyPipelines/Data/Big Analysis/all_var_AQ_h1.csv")
h1_data <- as.data.frame(h1_data)

# Software
h1_data$software[h1_data$software== 'eeglab_erplab'] <- 'eeglab'
h1_data$software[h1_data$software== 'eeglab_limo'] <- 'eeglab'
table(h1_data$software)

#Fill in empty responses
# High-pass filter type & direction
h1_data$ans_hf_type[h1_data$ans_hf_type== ''] <- 'unknown'
h1_data$ans_hf_direction[h1_data$ans_hf_direction == ''] <- 'unknown'
table(h1_data$ans_hf_type)
table(h1_data$ans_hf_direction) # has almost mo variability - remove later

# Segment exclusion criteria
h1_data$ans_exclusion_criteria_seg[h1_data$ans_exclusion_criteria_seg == ''] <- 'unknown'
table(h1_data$ans_exclusion_criteria_seg)

#Time window start, end & length
#exchange missing values to the mean of the column
h1_data$ans_time_w_start_h1[is.na(h1_data$ans_time_w_start_h1)] <- mean((h1_data$ans_time_w_start_h1), na.rm = T) 
h1_data$ans_time_w_end_h1[is.na(h1_data$ans_time_w_end_h1)] <- mean((h1_data$ans_time_w_end_h1), na.rm = T) 
h1_data$time_w_length_h1[is.na(h1_data$time_w_length_h1)] <- mean((h1_data$time_w_length_h1), na.rm = T) 
#table(h1_data$ans_time_w_start_h1)

# ICA algorythm
h1_data$ans_ica_algo[h1_data$ans_ica_algo == ''] <- 'unknown'
table(h1_data$ans_ica_algo)

#Baseline start and stop
h1_data$ans_baseline_start[h1_data$ans_baseline_start == '-200 ms'] <- '-200'
h1_data$ans_baseline_start[h1_data$ans_baseline_start == '-200ms'] <- '-200'
h1_data$ans_baseline_start <- as.numeric(h1_data$ans_baseline_start)
h1_data$ans_baseline_start[is.na(h1_data$ans_baseline_start)] <- mean((h1_data$ans_baseline_start), na.rm = T) 
h1_data$ans_baseline_stop[is.na(h1_data$ans_baseline_stop)] <- mean((h1_data$ans_baseline_stop), na.rm = T) 
```

## Prepare data for regression analysis

```{r}
# high collinearity problem removal
h1_data <- h1_data[-which(h1_data$mc_method_h1=='rft'),] # 3 teams removed

# too few samples that give Nan results in our models:
# ans_software_hostBESA Research 7.1  , ans_software_hostPython 3.8.10  , ans_hf_directionfwr_bwr, ans_ica_algoother
h1_data$ans_software_host[h1_data$ans_software_host == 'BESA Research 7.1'] <- 'unknown'
h1_data$ans_software_host[h1_data$ans_software_host == 'Python 3.8.10'] <- 'unknown'
h1_data$ans_software_host[h1_data$ans_software_host == 'R 4.1.3'] <- 'unknown'
h1_data$ans_hf_direction[h1_data$ans_hf_direction == 'fwr_bwr'] <- 'unknown'
h1_data$ans_ica_algo[h1_data$ans_ica_algo == 'other'] <- 'unknown'

# outliers removal
hf_out <- boxplot(h1_data$hf_cutoff)$out
h1_data$hf_cutoff[which(h1_data$hf_cutoff == hf_out)] <- mean((h1_data$hf_cutoff), na.rm = T) # removed one team that had 3 for a  HP filter

# cut data into dependent and independent variables
dependent_h1 <- h1_data[,c(27,28)] # binary H1 outcome and pvalues
h1_data <- h1_data[,c(2:26)] #delete the row numbers and dependent variables

# look if there are any NaN values
which(is.na(h1_data),arr.ind=TRUE) # row 157 column 15 is NA (Q: if team corrected for multiple comparisons 1 and 0)

# Caution: remove them from the dataset
dependent_h1 <- dependent_h1[-157,]
h1_data <- h1_data[-157,]

# Select only numeric predictors
mynumericdata <- h1_data %>%
  dplyr::select_if(is.numeric) 

shapiro.test(mynumericdata$hf_cutoff)#no data is normal
```

## Z-score p-values and normalize numeric data
```{r}

hist(qnorm(dependent_h1$pval/2, lower.tail = F))
pval_zscore <- qnorm(dependent_h1$pval/2, lower.tail = F)

indx_inf <- which(!is.finite(pval_zscore))
pval_zscore <- pval_zscore[-indx_inf] # remove the infinite value due to a mistakenly assigned 0 to pval it also takeS NAN values
h1_data <- h1_data[-indx_inf,] # remove those rows from independent variables too

continuous <- c(3, 7, 8, 10, 16:18, 23:25)

#normalizing data
for (i in 1:length(continuous)) {
  contin_data_col <- h1_data[,continuous[i]]
  h1_data[,continuous[i]] <- scale(contin_data_col)
}
```

## Build models
```{r}
#add pval
h1_data$zscore_p <- pval_zscore

# models including full data and but no interactions
model_complete <- glm(zscore_p ~ ., data = h1_data)

#summarize the model
summary(model_complete)

#Software
test_soft <- h1_data
test_soft$software <- as.factor(test_soft$software)
test_soft <- within(test_soft, software <- relevel(software, ref = 'eeglab'))
model_soft <- glm(zscore_p ~ software, data = test_soft)
summary(model_soft)

# Build reduced models including one predictor

models_list_single_1 = list()
for (g in 1:ncol(h1_data[,-1])){
  reduced_1 = glm(zscore_p ~ h1_data[,g], data = h1_data)
  models_list_single_1 = append(models_list_single_1, list(reduced_1))
}

```

## Compare models using AIC
The Akaike information criterion (AIC) is a mathematical method for evaluating how well a model fits the data it was generated from. In statistics, AIC is used to compare different possible models and determine which one is the best fit for the data.

Interpretation:

**K:** The number of parameters in the model. The default K is 2, so a model with one parameter will have a K of 2 + 1 = 3.

**AICc:** The information score of the model (the lower-case ‘c’ indicates that the value has been calculated from the AIC 
test corrected for small sample sizes). The smaller the AIC value, the better the model fit.

**Delta_AICc:** The difference in AIC score between the best model and the model being compared: In our data it looks like 
the first two models are equally good (model 15 and 23).

**AICcWt:** AICc weight, which is the proportion of the total amount of predictive power provided by the full set of models 
contained in the model being assessed. In this case, the top model contains 23% of the total explanation that can be found 
in the full set of models.

**Cum.Wt:** The sum of the AICc weights. The cumulative AICc weight is 0.46, meaning 46% of the support in the data are for these two models.
```{r}
#Put all models into the list:
models_list_full2 <- c(models_list_single_1, list(model_complete))

# Run AIC
aictab(cand.set = models_list_full2)

# Models 26 is the complete model

```

## Fit the best model acording to ACI
# Multiple comparissons (80 teams corrected for multiple comparissons, and 87 did not)
```{r}
model_mt <- glm(zscore_p ~ ans_mt_h1, data = h1_data)
summary(model_mt)

#Scatterplot
ggplot(h1_data, aes(x=zscore_p, y=ans_mt_h1)) +
  geom_point() 
```

```{r}
#boxplot of original pvalues to compare

library(ggplot2)
orig_data <- read.csv("C:/Users/ecesnait/Desktop/EEGManyPipelines/Data/Big Analysis/all_var_AQ_h1.csv")
orig_data <- as.data.frame(orig_data)
indx_no_mt <- which(orig_data$ans_mt_h1==F,arr.ind=TRUE)
indx_mt <- which(orig_data$ans_mt_h1==T,arr.ind=TRUE)
median(orig_data$pval[indx_no_mt], na.rm = T)
median(orig_data$pval[indx_mt], na.rm = T)

ggplot(orig_data, aes(x=as.factor(ans_mt_h1), y=pval)) + 
  geom_boxplot(fill="slateblue", alpha=0.2) + 
  xlab("multiple comparisons")+
  geom_text(x=1, y=1, label=median(orig_data$pval[indx_no_mt], na.rm = T), size = 3)+
  geom_text(x=2, y=1, label=median(orig_data$pval[indx_mt], na.rm = T), size = 3)#+ylim(min(orig_data$pval), 0.05)

cor.test(orig_data$pval, qnorm(orig_data$pval/2, lower.tail = F))
cor.test(orig_data$pval, qnorm(orig_data$pval/2, lower.tail = F))

```

# Baseline start
Interestingly, the start of the baseline window is linked to the size of a p-value and not the total length of the baseline window (Model26) or the end of it (mModel25)
```{r}
model_bs <- glm(zscore_p ~ ans_baseline_start, data = h1_data)
summary(model_bs)

#Scatterplot without scaling
h1_data <- read.csv("C:/Users/ecesnait/Desktop/EEGManyPipelines/Data/Big Analysis/all_var_AQ_h1.csv")
h1_data <- as.data.frame(h1_data)
#Baseline start and stop
h1_data$ans_baseline_start[h1_data$ans_baseline_start == '-200 ms'] <- '-200'
h1_data$ans_baseline_start[h1_data$ans_baseline_start == '-200ms'] <- '-200'
h1_data$ans_baseline_start <- as.numeric(h1_data$ans_baseline_start)
h1_data$ans_baseline_start[is.na(h1_data$ans_baseline_start)] <- mean((h1_data$ans_baseline_start), na.rm = T) 
h1_data$ans_baseline_stop[is.na(h1_data$ans_baseline_stop)] <- mean((h1_data$ans_baseline_stop), na.rm = T) 

ggplot(h1_data, aes(x=qnorm(pval), y=ans_baseline_start)) +
  geom_point() +
  geom_smooth(method=lm, se=FALSE)
```