---
title: "Logistic regression h1"
author: "Elena C"
date: "2023-11-20"
output: html_document
---


H1 states: *"There is an effect of scene category (i.e., a difference between images showing man-made vs. natural environments) on the amplitude of the N1 component, i.e. the first major negative EEG voltage deflection."*

The data table included in the analysis contains info on the software, high-pass cutoff values, reference, sampling rate, number of excluded subjects, baseline window length, topographic region, number of channels included in the region, multiple comparisons method, statistical method, and whether the finding was significant or not for the h1 testing. 

## Load packages
```{r, message=F}
rm(list = ls())
library(tidyverse)
library(broom)
library(readxl)

```
## Load data and inspect it for typos
```{r}
data <- read_excel("/Users/ecesnaite/Desktop/BuschLab/EEGManyPipelines/git/EEGManyPipelines-personal/Matlab/Big Analysis/initial_variables_h1_pval.xlsx")
data <- as.data.frame(data)
#correct values
table(data$software)
data$software[data$software== 'eeglab_erplab'] <- 'eeglab'
data$software[data$software== 'eeglab_limo'] <- 'eeglab'

data <- data[-which(data$mc_method_h1=='rft'),]#high multicollinearity with spm
#table(data$hf_cutoff)
#table(data$software)
#table(data$result_h1)
#hist(data$pval)
#hist(data$pval[data$pval<0.05])
```

## Inspect the data for outliers

Since these are real life preprocessing practices, should we consider possible outliers?

```{r, fig.show='hide'}
hf_out <- boxplot(data$hf_cutoff)$out
data$hf_cutoff[which(data$hf_cutoff == hf_out)] <- mean((data$hf_cutoff), na.rm = T)
```
## Inspect for normality
``` {r}
# Select only numeric predictors
data_reduced <- data[,c(-1,-13)] #delete the row numbers and binary reponse to hypothesis testing
data_reduced <- na.omit(data_reduced)
mynumericdata <- data_reduced %>%
  dplyr::select_if(is.numeric) 

shapiro.test(mynumericdata$hf_cutoff)#no data is normal

pval_zscore <- qnorm(data_reduced$pval)

indx_inf <- which(!is.finite(pval_zscore))
pval_zscore <- pval_zscore[-indx_inf] # remove the infinite value due to a mistakenly assigned 0 to pval
data_reduced <- data_reduced[-indx_inf,]

#normalise numeric data
for (i in 1:length(data_reduced)) {
  
  if (is.numeric(data_reduced[,i])){
    data_reduced[,i] <- data_reduced[,i]-mean(data_reduced[,i])/sd(data_reduced[,i])
  }
}
# scaled_data <- mutate_if(data, is.numeric, scale)
#table(scaled_data$pval)

data_reduced$pval <- pval_zscore
hist(data_reduced$pval)
#hist(data$hf_cutoff-mean(data$hf_cutoff)/sd(data$hf_cutoff))
```

## Check pvalues for true/false h0 testing
```{r}
# Boxplot basic
data %>%
  ggplot( aes(x=result_h1, y=pval, fill=result_h1)) +
  geom_boxplot() +
  theme(
    legend.position="none",
    plot.title = element_text(size=11)
  ) 

#zscored
data %>%
  ggplot( aes(x=result_h1, y=qnorm(pval), fill=result_h1)) +
  geom_boxplot() +
  theme(
    legend.position="none",
    plot.title = element_text(size=11)
  ) 
````

## Build a logistic model

```{r}

model <- glm(pval ~., data = data_reduced)

# Summarize the model
summary(model)
```

Multiple comparissons option 'none'  shows significant resutls (p<0.01). However, directionality is unexpected. Inspect visually this relationship.
``` {r}
#install.packages('hrbrthemes')
library(hrbrthemes)
library(viridis)
#inspect the relationship between mc method and zscore pvalues
data_reduced %>%
  ggplot( aes(x=mc_method_h1, y=pval, fill=mc_method_h1)) +
  geom_boxplot() +
  scale_fill_viridis(discrete = TRUE, alpha=0.6) +
  geom_jitter(color="black", size=0.4, alpha=0.9) +
  theme(legend.position="none", text = element_text(size=20),axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  labs(y= "z-scored p-val", x = "multiple comparissons")

#inspect the relationship between not using mc and original pvalues
mc_none <- data[data$mc_method_h1=='none',]
boxplot(mc_none$pval, ylab = 'original pval', xlab = 'mc none')
```