---
title: "log reg binary h1"
author: "Elena C"
date: "2023-11-29"
output: html_document
---

H1 states: *"There is an effect of scene category (i.e., a difference between images showing man-made vs. natural environments) on the amplitude of the N1 component, i.e. the first major negative EEG voltage deflection."*

The data table included in the analysis contains info on the software, high-pass cutoff values, reference, sampling rate, number of excluded subjects, baseline window length, topographic region, number of channels included in the region, multiple comparisons method, statistical method, and whether the finding was significant or not for the h1 testing. 

## Load packages
```{r, message=F}
rm(list = ls())
library(tidyverse)
library(broom)
library(readxl)

```
## Load data and inspect it for typos
```{r}
data <- read_excel("/Users/ecesnaite/Desktop/BuschLab/EEGManyPipelines/git/EEGManyPipelines-personal/Matlab/Big Analysis/initial_variables_h1_pval.xlsx")
data <- as.data.frame(data)
#correct values
table(data$software)
data$software[data$software== 'eeglab_erplab'] <- 'eeglab'
data$software[data$software== 'eeglab_limo'] <- 'eeglab'

data <- data[-which(data$mc_method_h1=='rft'),]#high multicollinearity with spm
#table(data$hf_cutoff)
#table(data$software)
#table(data$result_h1)
#hist(data$pval)
#hist(data$pval[data$pval<0.05])
```

## Inspect the data for outliers

Since these are real life preprocessing practices, should we consider possible outliers?

```{r, fig.show='hide'}
hf_out <- boxplot(data$hf_cutoff)$out
data$hf_cutoff[which(data$hf_cutoff == hf_out)] <- mean((data$hf_cutoff), na.rm = T)
```
## Inspect for normality
``` {r}
# Select only numeric predictors
data <- data[,c(-1,-14)] #delete the row numbers and pvalues of hypothesis testing
data <- na.omit(data)
mynumericdata <- data %>%
  dplyr::select_if(is.numeric) 

shapiro.test(mynumericdata$hf_cutoff)#no data is normal

#normalise numeric data
for (i in 1:length(data)) {
  
  if (is.numeric(data[,i])){
    data[,i] <- data[,i]-mean(data[,i])/sd(data[,i])
  }
}
# scaled_data <- mutate_if(data, is.numeric, scale)
#table(scaled_data$pval)
#hist(data$result_h1)
#hist(data$hf_cutoff-mean(data$hf_cutoff)/sd(data$hf_cutoff))
```


## Build a logistic model
``` {r}

model <- glm(result_h1 ~., data = data, family = 'binomial')

# Summarize the model
summary(model)
```

Time window of interest and baseline window length shows significant resutls (p<0.01). However, the effect seems to be influenced by a few extreme values.
``` {r}
# # plot the time window effect
data %>%
  mutate(prob = ifelse(result_h1 == "TRUE", 1, 0)) %>%
  ggplot(aes(time_window_interest_h1, prob)) +
  geom_point(alpha = 0.2) +
  geom_smooth(method = "glm", method.args = list(family = "binomial")) +
  labs(
    title = "Logistic Regression Model",
    x = "Time window",
    y = "Probability of confirming h1"
  )

# plot the baseline window effect
data %>%
  mutate(prob = ifelse(result_h1 == "TRUE", 1, 0)) %>%
  ggplot(aes(bs_window, prob)) +
  geom_point(alpha = 0.2) +
  geom_smooth(method = "glm", method.args = list(family = "binomial")) +
  labs(
    title = "Logistic Regression Model",
    x = "Baseline window",
    y = "Probability of confirming h1"
  )
```

## Assessing model accuracy
The model accuracy is measured as the proportion of observations that have been 
correctly classified. Inversely, the classification error is defined as the proportion 
of observations that have been misclassified.

``` {r}
# Make predictions
probabilities <- model %>% predict(data, type = "response")
head(probabilities)
```
# Logistic regression diagnostics
``` {r}
predictors <- colnames(mynumericdata)

# Bind the logit and tidying the data for plot
mynumericdata <- mynumericdata %>%
  mutate(logit = log(probabilities/(1-probabilities))) %>%
  gather(key = "predictors", value = "predictor.value", -logit)

#create scatterplots to inspect linearity
ggplot(mynumericdata, aes(logit, predictor.value))+
  geom_point(size = 0.5, alpha = 0.5) +
  geom_smooth(method = "loess") + 
  theme_bw() + 
  facet_wrap(~predictors, scales = "free_y") # doesn't look good

plot(model, which = 4, id.n = 3) # one extreme value that drags the effect

# Extract model results
model.data <- augment(model) %>% 
  mutate(index = 1:n()) 
```

# Check for outliers
``` {r}
outlier <- model.data %>% top_n(3, .cooksd) # The data for the top 3 largest values, according to the Cook’s distance
# plot standardized residuals
ggplot(model.data, aes(index, .std.resid)) + 
  geom_point(aes(color = result_h1), alpha = .5) +
  theme_bw()
model.data %>% 
  filter(abs(.std.resid) > 3)
```

# Colinearity
```{r}
library(car)
car::vif(model)#high colinearity

#inspect which variables are correlated
alias( model ) # software spm

vif_values <- vif(model)           #create vector of VIF values

barplot(vif_values, main = "VIF Values", horiz = TRUE, col = "steelblue") #create horizontal bar chart to display each VIF value

abline(v = 5, lwd = 3, lty = 2) #add vertical line at 5 as after 5 there is severe correlation

```